Fast and scalable Gaussian process modeling of stellar variability \vspace{-3\baselineskip}

### 2016-12-21 paper/ms.tex
Reran exp-squared fit using Sturm's theorem; updated paper

A stationary, smooth kernel may generally be expressed as a Taylor series for the $\tau \ne 0$ component,
plus the white noise component:
\begin{equation}
k(\tau) = a_0 \delta(\tau) +  \sum_{j=0}^\infty c_j \vert\tau\vert^j.
\end{equation}
Note that this encompasses all smooth, time-stationary kernels, and generally $c_1 \le 0$ as
correlated noise initially declines with separation in time.  We wish to approximate this
as the sum of damped cosine kernels, and since this general kernel is symmetric in $\tau$, we
can carry out the approximation in terms of sums of damped cosine kernel components
\begin{equation}
K_j(\tau,a_j,\beta_{R,j},\beta_{I,j}) = a_j \exp{(\beta_{R,j} \tau)}\cos{(\beta_{I,j} \tau)},
\end{equation}
which has a Taylor series expansion:
\begin{eqnarray}
\exp{(\beta_R \tau)}\cos{(\beta_I \tau)} &=& 1 - \beta_R \tau + \frac{1}{2}\left(\beta_R^2 - \beta_I^2\right)\tau^2 \\c &-&
\frac{1}{6} \left(\beta_R^3-3\beta_R\beta_I^2\right)\tau^3 + \frac{1}{24} \left(\beta_R^4+\beta_I^4-6\beta_R^2\beta_I^2\right) \tau^4+ O(\tau^5),
\end{eqnarray}
which may be continued to any order in $\tau$ as necessary.
If an approximation to $k(\tau)$ is desired in a Taylor series to some high order $N$, this
requires combining $N/2$ components of $K_j(\tau,a_j,\beta_{R,j},\beta_{I,j})$ with $\beta_{R,j}$ and $\beta_{I,j}$
chosen to match the coefficients $c_j$.  Note that for $N=1$ we only require a damped exponential
kernel with $\beta_I = 0$.

Another way to view this is in the power spectrum of the gaussian process, which is the
Fourier transform of the autocorrelation function.  The power spectrum of the $K_j$ kernel
is a Lorentzian, for which the frequency is given by $\omega = \beta_{I,j}$, while the
width of the Lorentzian is determined by $\beta_{R,j}$.  The functional form of this
power spectrum is: {\color{red} Check that this is true.}
\begin{equation}
P(\vert f\vert) = a_j \beta_{R,j} \frac{1}{\beta_{R,j}^2 + (2\pi f - \beta_{I,j})^2},
\end{equation}
where $f$ is the frequency in Hz.  The counterpart to the Taylor series is the observation
that a smooth distribution may be approximated arbitrarily precisely by summing together
Lorentzians with various values of $(a_j,\beta_{R,j},\beta_{I,j})$.  In the case that
a peak in the power spectrum is narrow at a specific frequency, then a small value of
$\beta_{R,j}$ gives a narrow Lorentzian that may approximate that peak.  If the wings
of the distribution decay faster than $f^{-2}$, then differences between Lorentzians
can be arranged to cancel in the wings, giving features in the power spectrum that have
narrower wings.

The functional form of $K_j$ may be used to approximate some commonly used Kernel components.
The squared exponential (Gaussian) kernel and the Mat\'ern kernels are commonly
used radial kernels, while the cosine kernel and the exponential-sine-squared kernels are
commonly used periodic kernels.  Linear combinations of our basis of damped cosine kernels may be
used to approximate each of these kernels to better than 1\% with various numbers of damped
cosine kernel components:
\begin{itemize}
\item {\bf Constant kernel.} A constant kernel component can be used to capture long-timescale
correlations in data, and can be represented exactly with $\beta_R = \beta_I =  0$.
\item {\bf Exponential kernel.} This kernel, $k(\tau) = e^{-\vert \tau/\tau_0\vert}$, is exactly
represented with $\beta_I=0$, $\beta_R = \tau_0^{-1}$.
\item {\bf Cosine kernel.} The cosine kernel, $k(t) = \cos{\left(\frac{2\pi}{P} \vert \tau \vert\right)}$,
may be represented exactly with $\beta_R=0$ and $\beta_I = \frac{2\pi}{P}$.
\item {\bf Squared exponential.} The squared exponential or Gaussian kernel, $G(z)=e^{-z^2/2}$,
has the behavior of declining steeply at large time separation, is an even function, and has $c_1$ = 0.
Likewise, the Fourier transform of the squared exponential kernel is $\hat G(\omega) = e^{-\omega^2/2}$,
which also declines steeply at large $\omega$;  this doesn't match the more gradual decline of a
Lorentzian.  Nevertheless, the squared exponential kernel may be well approximated by the sum of
three exponential kernels:
\begin{eqnarray}
%G(z) &\approx& 2.751807 e^{-1.531250 z} \cos{(0.341489 z)} \\
% &-& 0.354626 e^{-4.753504 z} \cos{(2.042572 z)}\\
% &-& 1.396834 e^{-1.796743 z} \cos{(1.764168 z)},
G(z) &\approx& 2.670143 e^{-1.500258 z} \cos{(0.346731 z)} \\
 &-& 0.458705 e^{-4.125588 z} \cos{(1.390912 z)}\\
 &-& 1.211438 e^{-1.744602 z} \cos{(1.798436 z)},
\end{eqnarray}
which has an accuracy of better than $<2.5\times 10^{-4}$ for all z.
Note that the last three coefficients depend on the first six as we have constrained
the first three terms of the Taylor series to match that of a Gaussian exactly, $1-z^2/2$.
% [2.6701433354886324,1.5002583976327974,-0.3467305042550739,-0.4587049531317026,4.125587583864702,1.3909117346609636] -1.2114383823569297 1.7446017341309372 1.7984359925567543
%[2.751807496353032,1.5312499408182438,-0.34148888180984543,-0.35462563655835033,4.753504292338051,2.0425716395901157,-1.3968344706423566,1.7967431129750422,-1.7641684073374964
\item {\bf Mat\'ern $p+1/2$ kernel.}  The Mat\'ern kernel is given by:
\begin{equation}
C_{p+1/2}(x) = \sigma^2 e^{-\sqrt{(2p+1)}x} \frac{\Gamma(p+1)}{\Gamma(2p+1)}
\sum_{i=0}^p \frac{(p+i)!}{i!(p-i)!} (\sqrt{(8p+4)}x)^{p-i},
\end{equation}
where $p$ is a positive integer.  In the limit $p\rightarrow \infty$, this becomes the
squared exponential kernel: $C_\infty(x) = \sigma^2 e^{-x^2/2}$.  But, for small
values of $p$, the Mat\'ern kernel may be approximated well by the sum of
$2p$ exponential kernels.  The exponential kernel is the Mat\'ern 1/2 kernel,
$C_{1/2}(x) = \sigma^2 e^{-x}$ ($p=0$), while
$p=1$ is relevant to stellar variability, which we describe next.
\item {\bf Mat\'ern 3/2 kernel.}  The Mat\'ern $3/2$ ($p=1$) kernel is given by:
\begin{equation}
K_{3/2}(\tau) = C_{3/2}(\tau/\tau_0) = (1+v)e^{-v},
\end{equation}
where $v = \sqrt{3}\vert \tau/\tau_0\vert$.  This kernel has a Fourier transform\footnote{Using the normalization convention:
$\hat f(\omega) = \frac{1}{\sqrt{2\pi}} \int_{-\infty}^\infty f(t) e^{i\omega t} dt$}  of
\begin{equation}
\hat K_{3/2}(\omega) = \frac{\tau_0 6^{3/2}}{\pi^{1/2}} (3+\tau_0^2\omega^2)^{-2},
\end{equation}
where $\omega = 2\pi f$.  This kernel may be well approximated by the sum of two exponential kernels:
\begin{equation}
K_{3/2}(\tau) \approx (1-b)e^{-v}+be^{-v\frac{b-1}{b}},
\end{equation}
where $b \gg 1$ is a dimensionless parameter; this approximation is exact in the limit $b \rightarrow \infty$.  For
large values of $b$, the maximum error of the approximation scales as $2/(e^2b) \approx 0.27/b$; for
$b=100$, the error in this approximation is $<0.3$\%.  Note that the two exponential terms
have opposite signs, which may lead to numerical instability for values of $b$ that
are large.

One advantage of the Mat\'ern 3/2 kernel is that it is smooth across
$\tau=0$ since its derivative is zero at $\tau=0$.  The exponential kernel {\it does not}
have this property, which leads to qualitatively different noise properties on small
timescales due to the sharpness of the exponential kernel at $\tau=0$.  The Mat\'ern 3/2 kernel
turns out to be similar to the observed properties of stellar activity and granulation, and thus the approximate
formula is simultaneously physically relevant and computationally convenient.
The power spectrum of a single exponential scales as $f^{-2}$ at high frequency,
while the Mat\'ern kernel scales as $f^{-4}$ which more closely matches the power
spectrum due to activity and granulation in the Sun \citep{2004A&A...414.1139A,2009A&A...495..979M}.
\item {\bf Mat\'ern 5/2 kernel.}  The Mat\'ern 5/2 kernel is given by
\begin{equation}
K_{5/2}(t) = C_{5/2}(y) = \sigma^2 (1+y+y^2/3)e^{-y},
\end{equation}
where $y = \sqrt{5}\vert \tau/\tau_0\vert$.  The Fourier transform is:
\begin{equation}
\hat K_{5/2}(\omega) = \frac{2\tau_0 10^{5/2}}{3\pi^{1/2}} (5+\tau_0^2\omega^2)^{-3}.
\end{equation}
This kernel may be approximated well by the sum of four exponential kernels
\begin{eqnarray}
K_{5/2}(\tau) &\approx& -147.6965573 e^{-1.0097610 y} + 85.8875034 e^{-1.0621734 y}\\
 &-&4.6540017 e^{-0.8353745 y} + 67.4630744 e^{-0.9160394 y},
\end{eqnarray}
which is accurate to $<4\times 10^{-6}$ for $y \le 10$.
The Fourier transform has a steeper dependence at large frequencies, $f^{-6}$, and as we are not aware of a stellar
power spectrum that requires a steeper spectrum than this, we forego approximating further Mat\'ern kernels.
\item {\bf Exponential-sine-square kernel.}  This kernel is a periodic, positive valued kernel
which has been used to model quasi-periodic stellar variability due to star spots.   It is
given by $E(w,\Gamma) = e^{-\Gamma \sin^2{w}}$ where $w = \frac{\pi}{P}\vert \tau \vert$, which may be
expanded as a series in $\cos{(2kw)} = \cos{\frac{2\pi}{P} k \vert \tau \vert)}$,
with $k \in \mathbb{N}$.  We have expanded this to $k=4$, and fit for the coefficients, giving:
\begin{equation}
E(w,\Gamma) \approx \sum_{k=0}^4 \left(c_{k,1}+c_{k,2} e^{-c_{k,4}\Gamma^{c_{k,3}}}\right) \cos{(2kw)},
\end{equation}
with $c_{0,1}=0.24999, c_{0,2}=0.752803, c_{0,3}=0.960477, c_{0,4}=0.643393$ for $k=0$,
$c_{1,1}=0.450851, c_{1,2}=-0.445338, c_{1,3}=0.960477, c_{1,4}= 1.18102$ for $k=1$,
$c_{2,1}=0.206593, c_{2,2}=-0.207361, c_{2,3}=1.65902, c_{2,4}= 0.210641$ for $k=2$,
$c_{3,1}=0.0728449, c_{3,2}=-0.0729381, c_{3,3}=2.43149, c_{3,4}= 0.0471224$ for $k=3$, and
$c_{4,1}=0.019204, c_{4,2}=-0.0192129, c_{4,3}=3.27977, c_{4,4}= 0.0114198$ for $k=4$.
This approximation is accurate to better than 0.6\% for $\Gamma < 3$, with a standard
deviation of 0.1\%.
\end{itemize}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=\textwidth]{./approximate_kernels/kernel_approx.pdf}
\caption{Approximation of various kernels with the sums of damped cosines.
(a). Kernels that may be expressed exactly: constant (dark grey), exponential (cyan),
cosine (olive), and damped cosine (salmon).
(b). Approximation of the squared-exponential (Gaussian) kernel (pink) with three
damped cosine kernels (magenta). (c). The approximation of the exponential-sine-squared
kernel with $\Gamma =1$ (blue) and $\Gamma = 3$ (green). (d). Two Mat\'ern kernels,
$K_{3/2}$ (yellow/brown) and $K_{5/2}$ (violet).  The Mat\'ern 3/2 kernel
is approximated with $b=100$. The exact kernels are solid, while approximate are dashed.}
\label{kernel_approx}
\end{center}
\end{figure}

A standard practice in constructing kernels for solving problems with Gaussian processes
is to add or multiply the preceding standard kernel components.  In the case of damped cosine
kernels, the product of two kernels is:
\begin{eqnarray}
K_j(\tau)K_k(\tau) &=&
\frac{1}{2} a_k a_j e^{-(\beta_{R,j}+\beta_{R,k}) \tau} \left(\cos{(\beta_{I,j}-\beta_{I,k})\tau}+\cos{(\beta_{I,j}+\beta_{I,k})\tau}\right)\\
&=& \frac{1}{2} \left(K_j(\tau,a_ja_k,\beta_{R,j}+\beta_{R,k},\beta_{I,j}-\beta_{I,k})\right.\\
&+&\left. K_j(\tau,a_ja_k,\beta_{R,j}+\beta_{R,k},\beta_{I,j}+\beta_{I,k})\right).
\end{eqnarray}
which is the sum of two damped cosine kernels.  Thus, only addition is necessary to
construct new kernels from damped cosine kernels.

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[scale=0.175]{./twoterm.pdf}
\hspace{0.2in}
\includegraphics[scale=1.0]{./colorcode.pdf}
\end{center}
\caption{Pictorial description of the extended sparse matrix where $N=10$, $p_0=0$, and $p=2$, following
the example shown in \citet{Ambikasaran:2015}.}
\figurelabel{matrix_structure}
\end{figure}

>>>>

%A stationary, smooth kernel may generally be expressed as a Taylor series for the $\tau \ne 0$ component,
%plus the white noise component:
%\begin{equation}
%k(\tau) = a_0 \delta(\tau) +  \sum_{j=0}^\infty c_j \tau^j.
%\end{equation}
%Note that this encompasses all smooth, time-stationary kernels, and generally $c_1 \le 0$ as
%correlated noise initially declines with separation in time.  We wish to approximate this
%as the sum of damped cosine kernels, and since this general kernel is symmetric in $\tau$, we
%can carry out the approximation in terms of sums of damped cosine kernel components
%\begin{equation}
%K_j(\tau,a_j,\beta_{R,j},\beta_{I,j}) = a_j \exp{(\beta_{R,j} \tau)}\cos{(\beta_{I,j} \tau)},
%\end{equation}
%which has a Taylor series expansion:
%\begin{eqnarray}
%\exp{(\beta_R \tau)}\cos{(\beta_I \tau)} &=& 1 - \beta_R \tau + \frac{1}{2}\left(\beta_R^2 - \beta_I^2\right)\tau^2 \\
%- \frac{1}{6} \left(\beta_R^3-3\beta_R\beta_I^2\right)\tau^3 &+& \frac{1}{24} \left(\beta_R^4+\beta_I^4-6\beta_R^2\beta_I^2\right) \tau^4+ O(\tau^5),
%\end{eqnarray}
%which may be continued to any order in $\tau$ as necessary.
%If an approximation to $k(\tau)$ is desired in a Taylor series to some high order $N$, this
%requires combining $N/2$ components of $K_j(\tau,a_j,\beta_{R,j},\beta_{I,j})$ with $\beta_{R,j}$ and $\beta_{I,j}$
%chosen to match the coefficients $c_j$.  Note that for $N=1$ we only require a damped exponential
%kernel with $\beta_I = 0$.
%
%Another way to view this is in the power spectrum of the gaussian process, which is the
%Fourier transform of the autocorrelation function.  The power spectrum of the $K_j$ kernel
%is a Lorentzian, for which the frequency is given by $\omega = \beta_{I,j}$, while the
%width of the Lorentzian is determined by $\beta_{R,j}$.  The functional form of this
%power spectrum is: {\color{red} Check that this is true.}
%\begin{equation}
%P(\vert f\vert) = a_j \beta_{R,j} \frac{1}{\beta_{R,j}^2 + (2\pi f - \beta_{I,j})^2},
%\end{equation}
%where $f$ is the frequency in Hz.  The counterpart to the Taylor series is the observation
%that a smooth distribution may be approximated arbitrarily precisely by summing together
%Lorentzians with various values of $(a_j,\beta_{R,j},\beta_{I,j})$.  In the case that
%a peak in the power spectrum is narrow at a specific frequency, then a small value of
%$\beta_{R,j}$ gives a narrow Lorentzian that may approximate that peak.  If the wings
%of the distribution decay faster than $f^{-2}$, then differences between Lorentzians
%can be arranged to cancel in the wings, giving features in the power spectrum that have
%narrower wings.

The functional form of $k_j$ may be used to approximate some commonly used Kernel components.
The squared exponential (Gaussian) kernel and the Mat\'ern kernels are commonly
used radial kernels, while the cosine kernel and the exponential-sine-squared kernels are
commonly used periodic kernels.  Linear combinations of our basis of damped cosine kernels may be
used to approximate each of these kernels to better than 1\% with various numbers of damped
cosine kernel components:
\begin{itemize}
\item {\bf Constant kernel.} A constant kernel component can be used to capture long-timescale
correlations in data, and can be represented exactly with $\beta_R = \beta_I =  0$.
\item {\bf Exponential kernel.} This kernel, $k(\tau) = e^{-\vert \tau/\tau_0\vert}$, is exactly
represented with $\beta_I=0$, $\beta_R = \tau_0^{-1}$.
\item {\bf Cosine kernel.} The cosine kernel, $k(t) = \cos{\left(\frac{2\pi}{P} \vert \tau \vert\right)}$,
may be represented exactly with $\beta_R=0$ and $\beta_I = \frac{2\pi}{P}$.
\item {\bf Squared exponential.} The squared exponential or Gaussian kernel, $G(z)=e^{-z^2/2}$,
has the behavior of declining steeply at large time separation, is an even function, and has $c_1$ = 0.
Likewise, the Fourier transform of the squared exponential kernel is $\hat G(\omega) = e^{-\omega^2/2}$,
which also declines steeply at large $\omega$;  this doesn't match the more gradual decline of a
Lorentzian.  Nevertheless, the squared exponential kernel may be well approximated by the sum of
four exponential kernels:
\begin{eqnarray}
G(z) &\approx& 63.119512 e^{-1.479579 z} -32.173615 e^{-1.593691 z} -29.425156 e^{-1.388929 z}\\
&+& e^{-1.383928 z} \left[-0.519196\cos{(1.883396 z)} +0.285478 \sin{(1.883396 z)}\right] \\
\end{eqnarray}
which has an accuracy of better than $<3\times 10^{-3}$ for all z, and the power spectrum has
been constrained to be positive.
\item {\bf Mat\'ern $p+1/2$ kernel.}  The Mat\'ern kernel is given by:
\begin{equation}
C_{p+1/2}(x) = \sigma^2 e^{-\sqrt{(2p+1)}x} \frac{\Gamma(p+1)}{\Gamma(2p+1)}
\sum_{i=0}^p \frac{(p+i)!}{i!(p-i)!} (\sqrt{(8p+4)}x)^{p-i},
\end{equation}
where $p$ is a positive integer.  In the limit $p\rightarrow \infty$, this becomes the
squared exponential kernel: $C_\infty(x) = \sigma^2 e^{-x^2/2}$.  But, for small
values of $p$, the Mat\'ern kernel may be approximated well by the sum of
$2p$ exponential kernels.  The exponential kernel is the Mat\'ern 1/2 kernel,
$C_{1/2}(x) = \sigma^2 e^{-x}$ ($p=0$), while
$p=1$ is relevant to stellar variability, which we describe next.
\item {\bf Mat\'ern 3/2 kernel.}  The Mat\'ern $3/2$ ($p=1$) kernel is given by:
\begin{equation}
k_{3/2}(\tau) = C_{3/2}(\tau/\tau_0) = (1+v)e^{-v},
\end{equation}
where $v = \sqrt{3}\vert \tau/\tau_0\vert$.  This kernel has a Fourier transform\footnote{Using the normalization convention:
$\hat f(\omega) = \frac{1}{\sqrt{2\pi}} \int_{-\infty}^\infty f(t) e^{i\omega t} dt$}  of
\begin{equation}
\hat k_{3/2}(\omega) = \frac{\tau_0 6^{3/2}}{\pi^{1/2}} (3+\tau_0^2\omega^2)^{-2},
\end{equation}
where $\omega = 2\pi f$.  This kernel may be well approximated by the sum of two exponential kernels:
\begin{equation}
k_{3/2}(\tau) \approx (1-b)e^{-v}+be^{-v\frac{b-1}{b}},
\end{equation}
where $b \gg 1$ is a dimensionless parameter; this approximation is exact in the limit $b \rightarrow \infty$.  For
large values of $b$, the maximum error of the approximation scales as $2/(e^2b) \approx 0.27/b$; for
$b=100$, the error in this approximation is $<0.3$\%.  Note that the two exponential terms
have opposite signs, which may lead to numerical instability for values of $b$ that
are large.

One advantage of the Mat\'ern 3/2 kernel is that it is smooth across
$\tau=0$ since its derivative is zero at $\tau=0$.  The exponential kernel {\it does not}
have this property, which leads to qualitatively different noise properties on small
timescales due to the sharpness of the exponential kernel at $\tau=0$.  The Mat\'ern 3/2 kernel
turns out to be similar to the observed properties of stellar activity and granulation, and thus the approximate
formula is simultaneously physically relevant and computationally convenient.
The power spectrum of a single exponential scales as $f^{-2}$ at high frequency,
while the Mat\'ern kernel scales as $f^{-4}$ which more closely matches the power
spectrum due to activity and granulation in the Sun \citep{2004A&A...414.1139A,2009A&A...495..979M}.
\item {\bf Mat\'ern 5/2 kernel.}  The Mat\'ern 5/2 kernel is given by
\begin{equation}
k_{5/2}(t) = C_{5/2}(y) = \sigma^2 (1+y+y^2/3)e^{-y},
\end{equation}
where $y = \sqrt{5}\vert \tau/\tau_0\vert$.  The Fourier transform is:
\begin{equation}
\hat k_{5/2}(\omega) = \frac{2\tau_0 10^{5/2}}{3\pi^{1/2}} (5+\tau_0^2\omega^2)^{-3}.
\end{equation}
This kernel may be approximated well by the sum of four exponential kernels
\begin{eqnarray}
k_{5/2}(\tau) &\approx& -147.6965573 e^{-1.0097610 y} + 85.8875034 e^{-1.0621734 y}\\
 &-&4.6540017 e^{-0.8353745 y} + 67.4630744 e^{-0.9160394 y},
\end{eqnarray}
which is accurate to $<4\times 10^{-6}$ for $y \le 10$.
The Fourier transform has a steeper dependence at large frequencies, $f^{-6}$, and as we are not aware of a stellar
power spectrum that requires a steeper spectrum than this, we forego approximating further Mat\'ern kernels.
\item {\bf Exponential-sine-square kernel.}  This kernel is a periodic, positive valued kernel
which has been used to model quasi-periodic stellar variability due to star spots.   It is
given by $E(w,\Gamma) = e^{-\Gamma \sin^2{w}}$ where $w = \frac{\pi}{P}\vert \tau \vert$, which may be
expanded as a series in $\cos{(2kw)} = \cos{\frac{2\pi}{P} k \vert \tau \vert)}$,
with $k \in \mathbb{N}$.  We have expanded this to $k=4$, and fit for the coefficients, giving:
\begin{equation}
E(w,\Gamma) \approx \sum_{k=0}^4 \left(c_{k,1}+c_{k,2} e^{-c_{k,4}\Gamma^{c_{k,3}}}\right) \cos{(2kw)},
\end{equation}
with $c_{0,1}=0.24999, c_{0,2}=0.752803, c_{0,3}=0.960477, c_{0,4}=0.643393$ for $k=0$,
$c_{1,1}=0.450851, c_{1,2}=-0.445338, c_{1,3}=0.960477, c_{1,4}= 1.18102$ for $k=1$,
$c_{2,1}=0.206593, c_{2,2}=-0.207361, c_{2,3}=1.65902, c_{2,4}= 0.210641$ for $k=2$,
$c_{3,1}=0.0728449, c_{3,2}=-0.0729381, c_{3,3}=2.43149, c_{3,4}= 0.0471224$ for $k=3$, and
$c_{4,1}=0.019204, c_{4,2}=-0.0192129, c_{4,3}=3.27977, c_{4,4}= 0.0114198$ for $k=4$.
This approximation is accurate to better than 0.6\% for $\Gamma < 3$, with a standard
deviation of 0.1\%.
\end{itemize}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=\textwidth]{./approximate_kernels/kernel_approx.pdf}
\caption{Approximation of various kernels with the sums of damped cosines.
(a). Kernels that may be expressed exactly: constant (dark grey), exponential (cyan),
cosine (olive), and damped cosine (salmon).
(b). Approximation of the squared-exponential (Gaussian) kernel (pink) with three
damped cosine kernels (magenta). (c). The approximation of the exponential-sine-squared
kernel with $\Gamma =1$ (blue) and $\Gamma = 3$ (green). (d). Two Mat\'ern kernels,
$k_{3/2}$ (yellow/brown) and $k_{5/2}$ (violet).  The Mat\'ern 3/2 kernel
is approximated with $b=100$. The exact kernels are solid, while approximate are dashed.}
\label{kernel_approx}
\end{center}
\end{figure}

A standard practice in constructing kernels for solving problems with Gaussian processes
is to add or multiply the preceding standard kernel components.  In the case of damped cosine
kernels, the product of two kernels is:
\begin{eqnarray}
k_j(\tau)k_k(\tau) &=&
\frac{1}{2} a_k a_j e^{-(\beta_{R,j}+\beta_{R,k}) \tau} \left(\cos{(\beta_{I,j}-\beta_{I,k})\tau}+\cos{(\beta_{I,j}+\beta_{I,k})\tau}\right)\\
&=& \frac{1}{2} \left(k_j(\tau,a_ja_k,\beta_{R,j}+\beta_{R,k},\beta_{I,j}-\beta_{I,k})\right.\\
&+&\left. k_j(\tau,a_ja_k,\beta_{R,j}+\beta_{R,k},\beta_{I,j}+\beta_{I,k})\right).
\end{eqnarray}
which is the sum of two damped cosine kernels.  Thus, only addition is necessary to
construct new kernels from damped cosine kernels.

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[scale=0.175]{./twoterm.pdf}
\hspace{0.2in}
\includegraphics[scale=0.85]{./colorcode.pdf}
\end{center}
\caption{Pictorial description of the extended sparse matrix where $N=10$, $p_0=0$, and $p=2$, following
the example shown in \citet{Ambikasaran:2015}.}
\figurelabel{matrix_structure}
\end{figure}
